Structured Synchronous Reactive Programming for Game Development \\ \Large{Case Study: On Rewriting Pingus from C++ to \CEU}

### 2017-05-22 paper.tex
...

TODO: OO state + methods

@@

In this work, we advocate structured synchronous reactive programming as an
expressive and productive alternative for game logic development.
We present a case study of rewriting Pingus from C++ to \CEU with the
contributions as follows:
%
\begin{itemize}
\item Identification of five recurrent control-flow patterns that likely apply
      to other games.
\item Idiomatic code in \CEU for six selected behaviors in the game logic.
\item An in-depth qualitative analysis of the implementations in \CEU in
      comparison to C++.
\end{itemize}

TODO: falar mais sobre as 3 contribuicoes:
- control-flow patterns
    - o que eh isso
        - definir controle
    - como eh diferente de GoF
        - behavioral patterns
- ideia geral das solucoes
    - eliminar estados explicitos com estruturas de controle apropriadas
- o que eh analise qualitativa
    - como medimos

The rest of the paper is organized as follows:
Section~\ref{sec.codebase} gives an overview of the Pingus codebases in C++ and
\CEU.
Section~\ref{sec.patterns} describes the five control-flow patterns and
discusses six case studies.
Section~\ref{sec.related} discusses related work.
Section~\ref{sec.conclusion} concludes the paper.

@@

This work focuses on a qualitative analysis for the programming techniques
that we applied during the rewriting process.
Not all techniques result in reduction in LoCs (especially considering the
verbose syntax of \CEU), but have other effects such as eliminating shared
variables and dependencies between classes.
%<!--, helping on encapsulation and cohesion.-->
Nonetheless, the lines with lower ratio numbers above correlate to the parts of
the game logic that we consider more susceptible to structured reactive
programming.
For instance, the \emph{Pingu} behavior (\emph{ratio 0.80}) contains complex
animations that are affected by timers, game rules, and user interaction.
In contrast, the \emph{Option screen} (\emph{ratio 0.97}) is a simple UI grid
with trivial mouse interactions.

@@

The rewriting process consisted of identifying sets of callbacks implementing
\emph{control-flow behaviors} in the game and translating them to \CEU using
appropriate structured constructs.
As an example, a double mouse click is characterized by a first click, followed
by a maximum amount of time, followed by a second click.
This behavior depends on different events (clicks and timers) which have to
occur in a particular order.
In C++, the implementation involves callbacks crossing reactions to successive
events which manipulate state variables explicitly.

@@

\begin{enumerate}
\item \emph{Finite State Machines}:
    Event occurrences map to transitions between states that trigger
    appropriate actions comprising the behavior of a game entity.
\item \emph{Continuation Passing}:
    The completion of a long-lasting activity may carry a continuation, i.e.,
    some action to execute next in the game.
\item \emph{Dispatching Hierarchies}:
    Entities form a dispatching hierarchy in which a container that receives a
    stimulus automatically forwards it to its managed children.
\item \emph{Lifespan Hierarchies}:
    Entities form a lifespan hierarchy in which a terminating container entity
    automatically destroys its managed children.
\item \emph{Signaling Mechanisms}:
    Entities often need to communicate explicitly through signaling mechanisms,
    especially if there is no hierarchy relationship between them.
\end{enumerate}

@@

@@

TODO: non reactive, C++ integration

>>>>

@@

In this work, we advocate structured synchronous reactive programming as an
expressive and productive alternative for game logic development.
We present a case study of rewriting Pingus from C++ to \CEU with the
contributions as follows:
%
\begin{itemize}
\item Identifying five recurrent control-flow patterns that likely apply to
      other games:
        \emph{Finite State Machines},
        \emph{Continuation Passing},
        \emph{Dispatching Hierarchies},
        \emph{Lifespan Hierarchies},
        \emph{Signaling Mechanisms}.
\item Applying idiomatic code in \CEU as alternative solutions for six selected
      behaviors in the game logic.
\item Presenting an in-depth qualitative analysis of the proposed solutions in
      comparison to the original implementations in C++.
\end{itemize}
%
A control-flow pattern is a recurring technique to describe dependencies and
explicit orders between statements (or groups of statements) in a program.
For instance, consider how a key press stimulus propagates through the game
entities and also what happens with them if the stimulus causes the end of the
game.
%
\CEU supports primitives that help describing complex control-flow
relationships in the game logic more concisely.
%
The rewriting process consisted of identifying sets of callbacks implementing
control flow in the game and translating them to \CEU using appropriate
structured constructs.
%
As an example, a double mouse click is characterized by a first click, followed
by a maximum amount of time, followed by a second click.
This behavior depends on different events (clicks and timers) which have to
occur in a particular order.
In C++, the implementation involves callbacks crossing reactions to successive
events which manipulate state variables explicitly.
%
Although we documented in detail two use cases for each of the five patterns,
in this work we only present six cases due to space constraints.
%
This work focuses on a qualitative analysis for the programming techniques
that we applied during the rewriting process.
Not all techniques result in reduction in LoCs (especially considering the
verbose syntax of \CEU), but have other effects such as eliminating shared
variables and dependencies between classes.

The rest of the paper is organized as follows:
Section~\ref{sec.codebase} gives an overview of the Pingus codebases in C++ and
\CEU.
Section~\ref{sec.patterns} describes the five control-flow patterns and
discusses six case studies in detail.
Section~\ref{sec.related} discusses related work.
Section~\ref{sec.conclusion} concludes the paper.

@@

%%%
This work focuses on a qualitative analysis for the programming techniques
that we applied during the rewriting process.
Not all techniques result in reduction in LoCs (especially considering the
verbose syntax of \CEU), but have other effects such as eliminating shared
variables and dependencies between classes.
%<!--, helping on encapsulation and cohesion.-->
Nonetheless, the lines with lower ratio numbers above correlate to the parts of
the game logic that we consider more susceptible to structured reactive
programming.
For instance, the \emph{Pingu} behavior (\emph{ratio 0.80}) contains complex
animations that are affected by timers, game rules, and user interaction.
In contrast, the \emph{Option screen} (\emph{ratio 0.97}) is a simple UI grid
with trivial mouse interactions.
%%%

@@

use cases first -> patterns later

@@

\begin{enumerate}
\item \emph{Finite State Machines}:
    Event occurrences map to transitions between states that trigger
    appropriate actions comprising the behavior of a game entity.
\item \emph{Continuation Passing}:
    The completion of a long-lasting activity may carry a continuation, i.e.,
    some action to execute next in the game.
\item \emph{Dispatching Hierarchies}:
    Entities form a dispatching hierarchy in which a container that receives a
    stimulus automatically forwards it to its managed children.
\item \emph{Lifespan Hierarchies}:
    Entities form a lifespan hierarchy in which a terminating container entity
    automatically destroys its managed children.
\item \emph{Signaling Mechanisms}:
    Entities often need to communicate and affect each other explicitly through
    signaling mechanisms, especially if there is no hierarchy relationship
    between them.
\end{enumerate}

@@

    - como eh diferente de GoF
        - not much unlike behavioral patterns

@@

TODO: non reactive, C++ integration
- TODO: OO state + methods
- eliminar estados explicitos com estruturas de controle apropriadas
